Generation of TypeScript Declaration Files from JavaScript Code

### 2021-05-21 mplr-2021/programming21-reviews-answered.txt

> ----------------------- REVIEW 1 ---------------------
> SUBMISSION: 28
> TITLE: Generation of TypeScript Declaration Files from JavaScript Code
> AUTHORS: Fernando Cristiani and Peter Thiemann
> 
> ----------- Overall evaluation -----------
> SCORE: -1 (major revision)
> ----- TEXT:
> Automatically generating TypeScript type signatures is a worthwhile
> goal; a mechanical process for generating such declarations would be
> less tedious, less error-prone, and more scalable. Pursuing a run-time
> trace-based technique is also compelling, given the significant
> difficulties that dynamically typed languages pose for static
> analysis. The initial results of the tool are promising: 244
> TypeScript declarations were automatically generated from existing
> module definitions, and 70 of these could be compared and validated
> against manually written declarations.
> 
> However, I believe the paper falls short both in terms of technical
> formulation as well as evidence of practical utility.
> 
> With respect to technical approach, the paper says (p.3/4) that the
> use of examples and dynamic analysis, rather than static analysis, is
> a novelty of the proposed tool. But a similar approach has been
> pursued previously:
> 
> * Trace Typing: An Approach for Evaluating Retrofitted Type Systems (Andreasen et al. ECOOP 2016)
> 
> * Just-in-Time Static Type Checking for Dynamic Languages (Ren et al. PLDI 2016)
> 
> * Dynamic Inference of Static Types for Ruby (An et al. POPL 2011)
> 
> * JSTrace: Run-time Type-Discovery for JavaScript (Saftoiu et al. 2010, > http://cs.brown.edu/research/plt/dl/flowtyping/v1/saftoiu-typediscovery-thesis.pdf)
> 
> How do the proposed techniques compare? Among these, trace typing is
> particularly relevant: like the current paper, it uses Jalangi for
> instrumentation.

Added discussion to the related work section.

> Besides not discussing this related work, the descriptions in Sections
> 4.2 and the end of 4.3 are rather informal: What exactly is the subset
> of JavaScript/TypeScript that is supported?

ECMAScript 5.1

> What is the instrumented
> evaluator computing? How are types --- which include unions and
> intersections (at least for function types) --- being computed from
> the run-time information? The lack of precise details makes it
> difficult to understand what the tool currently does, how plausible it
> might be to extend to support more of JavaScript/TypeScript in the
> future, and how it relates the related work described above.

Added some details about type computation at end of section 4.

> With respect to the implemented tool, it is not clear how close it is
> to being useful in practice. Many TypeScript features, as described in
> Table 1 and in the Evaluation, are not yet supported. For the 244
> modules for which declarations could be generated (and for the 70
> which could be compared to existing manual declarations): How many
> lines of code are in the modules? How many definitions are exported?
> How many examples were provided? How many examples were added to the
> 10 modules that were marked solvable? And how do these characteristics
> compare to the remaining Definitely Typed modules not currently
> supported? Especially because of the questions about technical
> formulation and novelty, it seems important for the system to come
> with a more comprehensive practical evaluation.
> 
> Overall, there is a lot of potential for this work --- especially on
> the practical side, for building a useful system --- but I think the
> results are preliminary to publish in their current form.
> 
> 
> *** Additional Comments and Questions ***
> 
> Fig 1: Does the tool generate declarations like this one? (If so, where
> does the name l__opts come from?)

@@

> Fig 2: I spent some time wondering how the module names in (b) and (f)
> were chosen (the former appears to be camel case version of the module
> name, and the latter follows the name in the example program). Some of
> this explanation comes later, but perhaps it's worth adding more
> explanation here as this seems to be the primary place in the paper
> where the different kind of templates are explained.

Said that for Globtoregexp.I__opts already!

> Fig 3: Why not depict the example extraction step --- which sometimes
> includes manual addition of examples --- here? And consider
> representing Jalangi's role in this architecture.

Fernando?
F: Added new images:
  - mplr-2021/graphics/dts-generate-block-diagram-with-code-examples-block.pdf
  - mplr-2021/graphics/dts-generate-block-diagram-with-jalangi.pdf

### 2021-05-21 mplr-2021/main.tex
computing types from run-time information

As \texttt{dts-generate} is based on run-time information, we need code
fragments that import and use the JavaScript library and we need to be
able to obtain this run-time information while running the code.
We explain in Section~\ref{sec:initial-code-base} how we obtain code
fragments. 
To obtain run-time information, we instrument
the code fragments and the entire code base of the library with
Jalangi \cite{DBLP:conf/sigsoft/SenKBG13}. We extend Jalangi's
analysis modules to gather data flow and type information as required
for our purposes and explained in Section~\ref{sec:run-time-information}.
% Jalangi provides several
% analysis modules that we extended as needed to retrieve the required
% run-time information.
%, which is then saved in a JSON file.

@@

Finally, interfaces are created by exploring \texttt{getField} and \texttt{methodCall} interactions from the run-time information. We
gather the interactions for a specific argument and build the
interface by incrementally adding new properties. Interactions within
the \texttt{followingInteractions} field are recursively traversed,
building a new interface at each level up to a predefined depth. Interfaces that turn out
to be equivalent are merged.

>>>>

As \texttt{dts-generate} is based on run-time information, we need code
fragments that import and use the JavaScript library and we need to be
able to obtain this run-time information while running the code.
We explain in Section~\ref{sec:initial-code-base} how we obtain code
fragments. 
To obtain run-time information, we instrument
the code fragments and the entire code base of the library with
Jalangi2 \cite{DBLP:conf/sigsoft/SenKBG13}. We extend Jalangi's
analysis modules to gather data flow and type information as required
for our purposes and explained in
Section~\ref{sec:run-time-information}.
\texttt{dts-generate} supports full ECMAScript~5.1, which is
the latest JavaScript version supported by Jalangi2. 
% Jalangi provides several
% analysis modules that we extended as needed to retrieve the required
% run-time information.
%, which is then saved in a JSON file.

@@

Interfaces are created by exploring \texttt{getField} and \texttt{methodCall} interactions from the run-time information. We
gather the interactions for a specific argument and build the
interface by incrementally adding new properties. Interactions within
the \texttt{followingInteractions} field are recursively traversed,
building a new interface at each level up to a predefined depth. Interfaces that turn out
to be equivalent are merged.

To construct the type signature for a function, we inspect the
primitive types of each argument as well as the interface created from
interactions on the argument in the function body. We do the same for
the return type of the function. This procedure yields a candidate
type for each function call observed during the execution.
For instance, for the \lstinline{globToRegExp} examples shown in
Listing~\ref{code:code-example-extracted-motivating-example} we obtain
the following observed candidate types
\begin{verbatim}
globToRegExp : (glob: string, opts: undefined) -> RegExp
globToRegExp : (glob: string, 
                opts: { extended : bool,
                        globstar?: any,
                        flags?: any }) -> RegExp
\end{verbatim}
As the type for \lstinline{glob} and the return type
\lstinline{RegExp} are the same, we collapse the type for
\lstinline{opts} into a union type. The union with the
\lstinline{undefined} type is special and converted into an optional
argument so we arrive at a single collapsed candidate type
\begin{verbatim}
globToRegExp : (glob: string, 
                opts?: { extended : bool,
                         globstar?: any,
                         flags?: any }) -> RegExp
\end{verbatim}
We proceed in this way with merging further candidate types until we
reach a set of non-mergable types for the function. If this set
contains more than one candidate type, then we generate an overloaded
definition. For example, the candidate types
\begin{verbatim}
f(string): string
f(number): number
f(boolean): boolean
\end{verbatim}
cannot be merged as the return types are different. They give rise to
an overloaded definition with three alternatives. However, given the candidate types
\begin{verbatim}
g(string): string
g(number): string
g(boolean): boolean
\end{verbatim}
we can merge the first two types on the first argument (because all
remaining arguments and the return types are equal) by collecting \lstinline{string} and
\lstinline{number} into a union type. The resulting set of non-mergable types
is thus
\begin{verbatim}
g(string|number): string
g(boolean): boolean
\end{verbatim}
